---
output: 
    github_document:
        html_preview: false
---

# `{rastro}`
`{rastro}` is an attempt to explore the new type system provided by the `{vctrs}` package. In `{rastro}` I create some simple data containers for simple scenarios that I encounter in my scientific work.

Consider the following example. When an observation is carried out, obtained value is usually accompanied by an error. The error can be symmetric or asymmetric. This can be represented in a `tibble` using two or three columns. Alternatively, a special type can be created that stores observations and errors and can be easily transformed to and from `data.frame`.
```{r}
library(rastro)
library(tibble, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)

# Create an observation from a standalone value
new_obs(5)

# Add an optional error
new_obs(rnorm(5), runif(5, 0, 0.5))

# And asymmetric errors
new_obs(1:10, p_err = 2, n_err = 1)
```

With proper methods defined, `rastro_obs` can be used within `tibble`:
```{r}
tibble(Time = 1:10, Value = new_obs(rnorm(10), runif(10, 0.1, 0.5)))
```

`rastro_obs<T>` behaves like a generic of type `T`. Type conversion rules limit possible values of `T`, but this limitation can be lifted by simply implementing additional type-conversion rules.
```{r, error = TRUE}
# An integer observation
new_obs(5L)

# Strings are not allowed
new_obs("a")
```

Arithmetic can be also implemented. A simple scenario would be to multiply/divide an observed quantity by a scalar, which scales both observation and its error:
```{r}
new_obs(1:10, 4) * 0.5
```

Another useful case is manipulation of multiple observed quantities. For instance, subtracting some background/baseline estimate (which itself is not known precisely) is a common task. To get the error on the derived quantity, a simple error propagation rule can be applied. The application of such rule can be encoded into the arithmetic functions for such type:
```{r}
new_obs(30, 3) - new_obs(10, 4)
```

The result is `20 ± 5`. If the function is
<img src="https://latex.codecogs.com/gif.latex?f(x,y)&space;=&space;x&space;-&space;y" title="f(x,y) = x - y" />
Then 
<img src="https://latex.codecogs.com/gif.latex?\sigma_f&space;=&space;\sqrt{\left(\dfrac{\partial&space;f}{\partial&space;x}\sigma_x\right)^2&space;&plus;&space;\left(\dfrac{\partial&space;f}{\partial&space;y}\sigma_y\right)^2}" title="\sigma_f = \sqrt{\left(\dfrac{\partial f}{\partial x}\sigma_x\right)^2 + \left(\dfrac{\partial f}{\partial y}\sigma_y\right)^2}" />,
which is simply 
<img src="https://latex.codecogs.com/gif.latex?\sqrt{\sigma_x^2&space;&plus;&space;\sigma_y^2}" title="\sqrt{\sigma_x^2 + \sigma_y^2}" />
for a subtraction.

## A more complex scenario
A similar type can be devised for a physical quantity. I chose flux (energy per unit area and unit time). Flux measurements are associated with a "filter" and a zero-point flux value (necessary for conversion). In general, it is illegal to combine fluxes obtained in different filters. 
```{r}
new_flux(100, "B", "mJy")
```
This gives us flux of `100` milliJansky measured in filter `B`. If the error on this measurement is, say, `5 mJy`, it is possible to store it as
```{r}
new_obs(new_flux(100, "B", "mJy"), new_flux(5, "B", "mJy"))
```
In some cases the type inference can kick in and the error can be simply input as a `double`. 
This system prevents operation on fluxes obtained from different filters:
```{r, error = TRUE}
new_obs(new_flux(100, "B")) - new_obs(new_flux(50, "R"))
```
But allows operations for the same filter (but not within `rastro_obs<T>`, because it requires `T` to be squared and then to extract square root from the sum)
```{r}
new_flux(100, "B") - new_flux(50, "B")
```

## Some more utilities
Angle measurements can be tricky, especially if angles are measured in degrees. Here is a solution
```{r}
new_degr(c(370, -370))
```
Values are mapped to `[0;360]` interval. Base function support can be also defined:
```{r}
mean(new_degr(10:30))
```
And, most importantly,
```{r}
sin(new_degr(30))
cos(new_degr(60))
```

**Work is still in progress**